---
title: "Scandinavia"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("utils.R")
source("get_data.R")
scandic<-c("Norway", "Sweden", "Denmark", "Finland")
scandic_nor<-c("Norge", "Sverige", "Danmark", "Finland")


scandic_country_conf<-ncovConfirmedTidy %>% 
    filter(Country.Region %in% scandic)

scandic_country_dead<-ncovDeathsTidy %>% 
    filter(Country.Region %in% scandic)
scandic_country_rec<-ncovRecoveredTidy %>% 
    filter(Country.Region %in% scandic)
```

This document is a daily report of the Covid-19 situation Globaly.
The data used to produce this report is the John Hopkins dataset from github:
"https://github.com/CSSEGISandData/COVID-19"

## Total Overview `r today`
Number of recorded cases of covid-19 




## Confirmed cases of covid 19 


New records of cases pr capita
```{r new cases confirmed pop,fig.cap = " Top 15 countries with most new cases pr capita of covid-19", echo=FALSE,warning=FALSE,message=FALSE}

new_cases_pop(today,scandic_country_conf,"Confirmed")

```


## Fatalities related to covid-19

Reported fatalities related to Covid-19, last 24 hours, pr capita
```{r new cases fatalities pop,fig.cap = " Top 15 countries with most new fatalities pr capita of covid-19", echo=FALSE,warning=FALSE,message=FALSE}

new_cases_pop(today,scandic_country_dead,"dead")

```

## Sick and Dead curve

This plot displays a normalized graph of the the number of sick people by covid-19
sick=confirmed-dead-recovred
and the number of fatalities by date.
The plot shows top 20 most infected countries
```{r sick and dead,fig.cap = "Normalized sick and Dead", echo=FALSE,warning=FALSE,message=FALSE}
#Set todays date
sick_dead_date<-today


death_compare<-scandic_country_dead %>% 
  filter(value!=0) %>%
  group_by(Country.Region,date) %>%
  summarize(dead=sum(value)) 

recovered<-scandic_country_rec %>% 
  filter(value!=0) %>%
  group_by(Country.Region,date) %>%
  summarize(recovered=sum(value))


# Get top 20 most infected countries
countries_sick<- scandic_country_conf %>% 
  filter(date==sick_dead_date) %>% 
  group_by(Country.Region) %>% 
  summarize(total=sum(value)) %>% 
  arrange(desc(total)) %>% 
  head(n=20)

#plot_sick<-
  
  ncovConfirmedTidy %>% 
  filter(value!=0, Country.Region!="Cruise Ship", date<sick_dead_date) %>%
  group_by(Country.Region,date) %>%
  summarize(confirmed=sum(value)) %>% 
  left_join(death_compare) %>% 
  left_join(recovered) %>% 
  mutate(sick=confirmed-dead-recovered) %>%
  filter(sick>100) %>% 
  mutate(sick=normalize(sick), dead=normalize(dead)) %>% 
  pivot_longer(cols = c("confirmed","dead","recovered", "sick" ), names_to = "type", values_to = "value") %>% 
  mutate(value=replace_na(value,0)) %>% 
  filter(type!="confirmed",type!="recovered",Country.Region %in% scandic) %>% 
  ggplot(aes(x=date,y=value, color=type))+
  geom_line()+
  facet_wrap(~Country.Region, scales = "free")+
  theme_minimal()+
  labs(title="Sick and Dead-curve", y="Percentage")+
  scale_y_continuous(labels = scales::percent)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

#ggplotly(plot_sick)


```


## Confirmed and fatality curve comparison
A comparison of the confirmed cases of covid-19 and fatality curves for different countries. The


# Fatality curve
```{r Fatality curv,fig.cap = " Top 15 countries with most new fatalities pr capita of covid-19", echo=FALSE,warning=FALSE,message=FALSE}
#Pr capita dead, log
per_capita_dead<-ncovDeathsTidy  %>% 
  filter(value!=0, Country.Region%in% scandic) %>%
  group_by(Country.Region,date) %>%
  left_join(population_total) %>% 
  summarize(pr_date=sum(value)*100000/mean(population)) %>% 
  filter(pr_date>0.2)%>% 
  mutate(row_name = row_number())  %>% 
  ggplot(aes(x=row_name, y=pr_date, color=Country.Region))+
  geom_line()+
  scale_y_log10()+
  labs(x="Days after fatality of 0.02 pr 100000", y="Dead",color="Country", title = "Death Curve pr capita for different countries")

ggplotly(per_capita_dead)




``` 


## Overview pr Country `r today`
Number of recorded cases of covid-19 country

```{r total pr country, echo=FALSE,warning=FALSE,message=FALSE}

countries_total_raw<-"https://redutv-api.vg.no/corona/v1/world/reports?"%>% jsonlite::fromJSON() %>% simplify() 
countries_total<-as_tibble(countries_total_raw$data) %>% 
  mutate(Confimred=confirmed$current,
         Deaths=deaths$current) %>% 
  select(-confirmed,-deaths)


library(knitr)
library(kableExtra)
library(DT)
library(formattable)


customGreen0 = "#DeF7E9"

customGreen = "#71CA97"

customRed = "#ff7f7f"

countries_total %>%
  select(-isoCode,-population) %>%
  rename(Country=countryNorwegian)%>% 
  filter(Country %in% scandic_nor) %>% 
  drop_na() %>% 
  formattable(align =c("l","c","c","c", "r")
              ,
              list( Country = 
    formatter("span", style = x ~ formattable::style(color = "grey",font.weight = "bold")), 
  confirmedPer100kCapita = color_bar(customGreen),
  deathsPer100kCapita = color_bar(customRed)
  )
  ) %>% 
 as.datatable()



```

## Number of new confirmed and fatalities pr day

```{r new cases, echo=FALSE,warning=FALSE,message=FALSE}
# Time serie pr country ---------------------------------------------------
# Confirmed and dead pr country
pr_country_raw<-"https://redutv-api.vg.no/corona/v1/world/timeseries/by-country?"  %>% jsonlite::fromJSON()%>% simplify()
pr_country<-pr_country_raw$items

# Time serie of confirmed and dead pr country
pr_country_timeseries<-pr_country %>% select(name,timeseries) %>% unnest(cols = c(timeseries))




confirmed_countries<-pr_country_timeseries %>% 
  filter(date==tomorrow) %>%  
  arrange(desc(confirmed)) %>% 
  head(n=20)

dead_countries<-pr_country_timeseries %>% 
  filter(date==tomorrow) %>%  
  arrange(desc(deaths)) %>% 
  head(n=20)

  

pr_day<-pr_country_timeseries %>% 
    mutate(new_confirmed=new$confirmed,
         new_deaths=new$deaths,
         date=as_date(date)) 



confirmed_plot<-pr_day %>% 
  filter(name %in% scandic_nor) %>% 
  ggplot(aes(x=date, y=new_confirmed))+
    geom_col()+
    facet_wrap(~name, scales = "free")+
    geom_smooth()+  
    theme_minimal() + 
    scale_color_brewer(palette = "Dark2") +
    labs(title = "New Confirmed pr day", y="Number of Confirmed", x="")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

ggplotly(confirmed_plot)

dead_plot<-pr_day %>% 
  filter(name %in% scandic_nor) %>% 
    ggplot(aes(x=date, y=new_deaths))+
    geom_col()+
    facet_wrap(~name, scales = "free")+
    geom_smooth()+  
    theme_minimal() + 
    scale_color_brewer(palette = "Dark2") +
    labs(title = "New Fatalities pr day", y="Number of Fatalities", x="")+
      theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

ggplotly(dead_plot)
  
```